Acceleration of Derivative Calculations with Application to Radial Basis Function -- Finite-Differences on the Intel MIC Architecture

### 2014-01-17 acm/ieee_paper.tex
edits

\ee{The main idea of this paper, computing 16 derivatives at a time,
  can be applied on different architectures such as multiprocessor
  multicore or GPU. To illustrate this fact, we conducted similar
  experiments on a classical CPU architecture with two Intel Core i7
  (SandyBridge) processors, each has eight cores that support AVX
  instructions.}  Both the compiler-generated and the manually-written
vectorization using AVX instructions reach a (read) bandwidth of
70~GB/s. This indicates that the performance of the classical SpMV can
not overcome 17~Gflop/s. A generic implementation of the
multiplication of four vectors with four matrices on the {\tt RBF}
matrix reaches a performance of 27~Gflop/s. We also implemented an AVX
vectorized version which is less compact than the one on MIC because
of smaller vector registers and the abscence of a gather instruction
(AVX2 on Haswell processor has that feature). It reaches a
performance of 42~Gflop/s. This confirms that our method is sound and
generalizes to architectures other than the Intel MIC architecture.

@@

In future work, we will examine the effect of larger stencil sizes and
double precision. The presented performance does not reach the
predicted peak and we believe that common techniques from SpMV such as
cache partitioning can be successfully applied. We will also integrate
the techniques presented in this paper within a fluid simulation using
RBFFD. We will also look developping a similar technique on more
commonly available devices such as GPUs.

>>>>

\ee{The main idea of this paper, computing 16 derivatives at a time,
  can be applied on different architectures such as multiprocessor
  multicore or GPU. To illustrate this fact, we conducted similar
  experiments on a classical CPU architecture with two Intel Core i7
  (SandyBridge) processors, each has eight cores that support AVX
  instructions.}  Both the compiler-generated and the manually-written
vectorization using AVX instructions reach a (read) bandwidth of
70~GB/s. This indicates that the performance of the classical SpMV can
not overcome 17~Gflop/s. A generic implementation of the
multiplication of four vectors with four matrices on the {\tt RBF}
matrix reaches a performance of 27~Gflop/s. \ee{We also implemented an AVX
vectorized version which is less compact than the one on MIC because
of smaller vector registers and the abscence of a gather instruction
(Haswell processors support AVX2 which have that feature).} It reaches a
performance of 42~Gflop/s. This confirms that our method is sound and
generalizes to architectures other than the Intel MIC architecture.

@@

In future work, we will examine the effect of larger stencil sizes and
double precision. The presented performance does not reach the
predicted peak and we believe that common techniques from SpMV such as
cache partitioning can be successfully applied. We will also integrate
the techniques presented in this paper within a fluid simulation using
RBFFD. \ee{We will also look developping a similar technique on more
commonly available devices such as GPUs.}
